---
title: "R Notebook"
output: html_notebook
---

#-------------------------------------------------------------------------------
# 1.0. Data import
#-------------------------------------------------------------------------------

```{r}
library(readr)

df_v1 <- read_csv("Datasets/3_V1_processed.csv", 
    col_types = cols(id = col_character(), 
        end_timestamp_previous_ride = col_datetime(format = "%Y-%m-%d %H:%M:%S"), 
        start_timestamp = col_datetime(format = "%Y-%m-%d %H:%M:%S"), 
        zip_4 = col_double()))

df_v2 <- read_csv("Datasets/4_V2_processed.csv", 
    col_types = cols(id = col_character(), 
        end_timestamp_previous_ride = col_datetime(format = "%Y-%m-%d %H:%M:%S"), 
        start_timestamp = col_datetime(format = "%Y-%m-%d %H:%M:%S"), 
        zip_4 = col_double()))

df_v3 <- read_csv("Datasets/5_V3_processed.csv", 
    col_types = cols(id = col_character(), 
        end_timestamp_previous_ride = col_datetime(format = "%Y-%m-%d %H:%M:%S"), 
        start_timestamp = col_datetime(format = "%Y-%m-%d %H:%M:%S"), 
        zip_4 = col_double()))
```

#-------------------------------------------------------------------------------
# 2.0. Survival analysis preparation
#-------------------------------------------------------------------------------

```{r}
library(survival)
library(ranger)
library(ggplot2)
library(dplyr)
library(ggfortify)
library(skimr)
library(survminer)
library(pec) #SelectCox
```

# Experiment with groupoing of zip codes
```{r}
zip_4_wijknaam <- read_csv("External data/Zip_4_Wijknaam_v2.csv", 
    col_types = cols(...1 = col_skip(), Wijk2021 = col_skip()))

zip_4_wijknaam <- zip_4_wijknaam[-c(which(duplicated(zip_4_wijknaam$PC6))),] 
```

```{r}
df_v1 <- left_join(df_v1, zip_4_wijknaam, by = c("zip_6"="PC6"))
df_v2 <- left_join(df_v2, zip_4_wijknaam, by = c("zip_6"="PC6"))
df_v3 <- left_join(df_v3, zip_4_wijknaam, by = c("zip_6"="PC6"))
```


#-------------------------------------------------------------------------------
# 3.0. Cox Proportional Hazards Model
#-------------------------------------------------------------------------------

```{r}
options(scipen=999)
```

```{r}
df_v1$zip_4 <- as.factor(df_v1$zip_4)
df_v1$day <- as.factor(df_v1$day)
df_v1$month <- as.factor(df_v1$month)
df_v1$hour <- as.factor(df_v1$hour)
df_v1$hour <- relevel(df_v1$hour, ref = 13) #12:00 as baseline

df_v2$zip_4 <- as.factor(df_v2$zip_4)
df_v2$day <- as.factor(df_v2$day)
df_v2$month <- as.factor(df_v2$month)
df_v2$hour <- as.factor(df_v2$hour)
df_v2$hour <- relevel(df_v2$hour, ref = 13)

df_v3$zip_4 <- as.factor(df_v3$zip_4)
df_v3$day <- as.factor(df_v3$day)
df_v3$month <- as.factor(df_v3$month)
df_v3$hour <- as.factor(df_v3$hour)
df_v3$hour <- relevel(df_v3$hour, ref = 13)

df_v3$interval <- as.factor(df_v3$interval)
```

# Sun = 1, Mon = 2, Tue = 3, Wed = 4, Thu = 5, Fri = 6, Sat = 7

```{r}
df_v1$period <- "weekday"

for (i in 1:nrow(df_v1)){
  if (df_v1$day[i] == 1){ df_v1$period[i] <- "weekend"}
  if (df_v1$day[i] == 7){ df_v1$period[i] <- "weekend"}
}

df_v3$period <- "weekday"

for (i in 1:nrow(df_v3)){
  if (df_v3$day[i] == 1){ df_v3$period[i] <- "weekend"}
  if (df_v3$day[i] == 7){ df_v3$period[i] <- "weekend"}
}
```


#-------------------
# a) V1 
#-------------------

# Backward selection for Cox model
```{r}
selectCox(Surv(idle_time_in_hours, event_indicator) ~ 
                   tram_200+bus_200+metro_200+train_200+
                   playground_200+kindergarten_200+
                   college_200+school_200+university_200+
                   humidity+sunshine+precipitation+temperature+
                   hour+period+month+
                   zip_4, data = df_v1, rule = "aic")
```

# Run best model
```{r}
v1_all <- coxph(Surv(idle_time_in_hours, event_indicator) ~ 
                   tram_200+
                   playground_200+kindergarten_200+
                   school_200+
                   humidity+sunshine+precipitation+temperature+
                   hour+period+month+
                   zip_4
                   ,data = df_v1)
v1_all$concordance[6] # 0.5901 with period instead of day, #0.5840 with period and wijknaam2021
```

```{r}
summary(v1_all)
```

# Testing for time-varying coefficients
```{r}
zph_v1_all <- cox.zph(v1_all)
zph_v1_all #tram_200, playground_200, kindergarten_200, H+D+M, zip_4 are time-varying
```

#-------------------
# c) V3
#-------------------

# Cox model 3.0 based on previously relevant variables from V1 minus insignificant variables after first iteration (see #)
```{r}
v3_all <- coxph(Surv(start2, stop2, event_indicator) ~ 
                   #tram_200:strata(interval)+
                   playground_200:strata(interval)+kindergarten_200:strata(interval)+
                   #school_200+
                   #humidity+
                   sunshine+precipitation+#temperature+
                   hour+period+#month+
                   zip_4+cluster(id)
                   ,data = df_v3)
v3_all$concordance[6] #if dropoff instead of hour => 0.56, 
                      #if dropoff:strata(interval) instead of hour => 0.57 
                      #if period instead of day => 0.8383
                      #if period instead of day and wijknaam2021 instead of zip_4 => 0.8362
```

```{r}
summary(v3_all)
```

# Testing for time-varying coefficients 
```{r}
zph_v3_all <- cox.zph(v3_all)
zph_v3_all #kindergarten_200, H, zip_4
```

#-------------------------------------------------------------------------------
# 5.0. Other (irrelevant)
#-------------------------------------------------------------------------------

```{r}
zip_4_distr <- df_v1 %>%
  count(zip_4) %>%
  group_by(zip_4) %>%          # now required with changes to dplyr::count()
  mutate(prop = prop.table(n))

# Outliers: (3026, 3027, 3028, 3082, 3083, 3055, 3053) have very large coefficients, which is explained by the low proportion of trips in these areas (1,1,1,1,3)
```

```{r}
period_distr <- df_v1 %>%
  count(period) %>%
  group_by(period) %>%          # now required with changes to dplyr::count()
  mutate(prop = prop.table(n))

```

```{r}
df_v3$prec_dum <- 0

for (i in 1:nrow(df_v3)){
  if (df_v3$precipitation[i] > 2){
    df_v3$prec_dum[i] <- 1
  }
}
df_v3$prec_dum <- as.factor(df_v3$prec_dum)

df_v3$sun_dum <- 0

for (i in 1:nrow(df_v3)){
  if (df_v3$sunshine[i] > 50){
    df_v3$sun_dum[i] <- 1
  }
}
df_v3$sun_dum <- as.factor(df_v3$sun_dum)
```




```{r}
pc6hnr20210801_gwb$PC4 <- substr(pc6hnr20210801_gwb$PC6, 1, 4)
```

```{r}
pc6hnr20210801_gwb$check <- NA
pc6hnr20210801_gwb$PC4 <- as.numeric(pc6hnr20210801_gwb$PC4)

save <- subset(pc6hnr20210801_gwb, PC4 > 3010)
save <- subset(save, PC4 < 3084)  
save <- subset(save, select = c(PC6, Wijk2021))

pairs <- save[!duplicated(t(apply(save, 1, sort))),]
pairs$relevant <- NA


for (i in 1:nrow((pairs))){
  if (as.character(pairs$PC6[i]) %in% unique(as.character(df_v1$zip_6))){
    pairs$relevant[i] <- "keep"
  }
}

pairs <- subset(pairs, relevant == "keep")

check <- left_join(pairs, wijk2021, by = c("Wijk2021" = "wijkcode2021"))
check <- subset(check, select = c(PC6, Wijk2021, wijknaam2021))

write.csv(check,"Zip_4_Wijknaam_v2.csv", row.names = TRUE)
```

# playground
```{r}
png(file="time_uni.png",
width=1200, height=800)

op <- par(mar = c(7,8,4,2) + 0.5)
par(cex.axis=2, cex.lab=3)
plot(zph_v1_all_exp_time[6], lwd=6, ylim= c(-1.5,1.5), line = 5)
abline(0,0, col=1, lty=3, lwd=6)
#abline(h=v1_all_exp_time$coef[70], col=3, lwd=6, lty=2)
legend("bottomright", legend=c("Reference line for null effect", "Average hazard over time", "Time-varying hazard"), lty=c(3,2,1), col=c(1,3,1), lwd=6, cex=2)
par(op)

dev.off()
```








































